To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis

### 2020-04-05 Thesis.tex
Add files via upload
Thesis

	One of the first papers which highlights the issue from the point of view of Economics is work by \cite{avery1999market}. \cite{avery1999market} considered the influence of the recommender system on choice regarding the consumption of good for the sequentially arriving agents.

	\cite{bergemann2006optimal} address the question of optimal pricing under the presence of recommender system.

@@

	One of the first works that focus on design of optimal recommendation policy is the work by \cite{kremer2014}. The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff. One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher. The finite number of consumers arrive sequentially and each chooses between either of two alternatives. After the choice is made the consumer leaves forever. The platform accumulates reviews and commits to the recommendation policy it uses to make recommendations to the newcomers. \cite{kremer2014}, firstly, neatly in simple terms demonstrate why the full revelation policy may be suboptimal\footnote{Maybe the appearance of this paper and similar to that in 2014 was the reason why \textit{Amazon} decided to switch their recommendation policy from the fully revealing one.}. Secondly, the optimal recommender policy is derived which has a threshold nature. That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold. Otherwise, the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end.

@@

	Another paper, which is closely related to the one mentioned previously is the paper by \cite{li2018recommender} (actually there are two quite different versions of this paper). In the two papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.

@@

\begin{proposition}\label{notprecise}
	Given the commission rate $\alpha$ and precision $\beta$ the optimal recommender policy implies:
	$$\hat{x} = \frac{1}{2} - \frac{\gamma - \alpha}{\gamma \beta} \frac{p_A-p_B}{2} + \frac{\tilde{q}_A - \tilde{q}_B}{2 \beta}$$
	i.e. those customers for whom signal regarding their location falls to the left of $\hat{x}$ are recommended the seller $A$, otherwise they are recommended the seller $B$.
	Equilibrium prices are
	$$p_i = \frac{\gamma}{\gamma - \alpha} \beta + \frac{\gamma(\tilde{q}_i - \tilde{q}_{-i})}{3(\gamma - \alpha)}$$
	where
	$$\tilde{q}_i = \begin{cases}
	q_i, \text{ if firm }i\text{ discloses }\\
	\frac{q^*}{2}, \text{otherwise}
	\end{cases}$$
	and $$q^* = -\frac{6 \beta - 1}{3} + \frac{1}{3 \gamma} \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}$$
\end{proposition}
\begin{proof}
	Using \eqref{signal} and repeating the proof of the Proposition \ref{with} one can obtain the expression for $\hat{x}$ and equilibrium prices. $q^*$ is found in the same way as in the Proposition \ref{precise}.
\end{proof}
Note that for $\beta = 1$ i.e. for the highest precision of the recommender system, we get the exact result of the Proposition \ref{precise}. Several observations are worth noting from the Proposition above. First of all, low precision makes the impact of the recommender system less prominent. We have a multiplier $\frac{\gamma - \alpha}{2 \gamma \beta}$ before the term $(p_A - p_B)$ in the threshold which determines the demand for sellers. The lower is $\beta$ the stronger becomes incentives for the sellers to undercut the competitor. This can be explained in the following way: now there is a chance that the platform mistakes regarding the location of the customer, in this case, the platform makes a recommendation assuming that it is an average customer (location $\frac{1}{2}$). The latter increases the share of customers in the middle whom the sellers have the highest incentives to compete for. As a result, other things equal, equilibrium prices are lower, the less precise is the recommender system implemented by the platform.

@@

\begin{proposition}\label{profit_beta}
	\leavevmode 
	\begin{itemize}
		\item If $\beta \le 0.5$ and  $\gamma > 1$ then $\frac{\partial \pi }{\partial \alpha} > 0$ and the optimal commission is $\alpha = \bar{\alpha}$
		\item  If $\beta \le 0.5$ and $\gamma < 1$ then $\frac{\partial \pi }{\partial \alpha} < 0$ and the optimal commission is $\alpha = 0$	
		\item If $\beta > 0.5$ and  $\gamma > 1$ then $\frac{\partial \pi }{\partial \alpha} < 0$ and the optimal commission is $\alpha = 0$
		\item If $\beta > 0.5$ and $\gamma < 1$ then $\frac{\partial \pi }{\partial \alpha} > 0$ and the optimal commission is $\alpha = \bar{\alpha}$
	\end{itemize}
\end{proposition}
\begin{proof}
 First of all, we need to find the platform's expected profit. Given prices $p_A, p_B$ and qualities $q_A, q_B$, total benefit of the platform is equal to:
 \begin{align}\label{profit}
 \alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \biggl(\int_{0}^{\hat{x}} \biggl[ (\beta +(1-\beta)\hat{x})(V + q_A - p_A - x) + \nonumber \\+ (1-\beta)(1-\hat{x})(V + q_B - p_B - (1-x))\biggr]dx + \int_{\hat{x}}^1 \biggl[(1-\beta) \hat{x}(V+q_A-p_A-x) + \nonumber\\
 +(\beta + (1-\beta)(1-\hat{x}))(V + q_B - p_B -(1-x)) \biggr] dx\biggr) = \\
 = 	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(V - \frac{1}{2}+(1-\hat{x})(q_B - p_B) + \hat{x}(q_A-p_A) + \beta \hat{x} - \beta \hat{x}^2  \right)\nonumber
 \end{align}
 Where $\hat{x}$ is defined according to Proposition \ref{notprecise}. For those customers located to the left of $\hat{x}$ the probability of being recommended a seller $A$ is $\beta + (1-\beta) \hat{x}$. Since with probability $\beta$ the platform receives the true location of customer and with probability $(1-\beta) \hat{x}$ the platform receives a signal from the interval $[0, \hat{x}]$. In the following way, one can obtain probabilities of being recommended seller $B$ for those customers located to the left of $\hat{x}$ as well as probabilities of being recommended sellers $A$ and $B$ for the customers located to the right of $\hat{x}$. Finally, we obtain an expression in \eqref{profit}. Then, in the same way as in the Proposition \ref{opt_alpha} one can obtain the expression for the profit:
 \begin{align*}
 \pi = \frac{\gamma(54 \beta^2(1+4V) - (1-q^{*3})(1-2\beta) - 216 \beta^3)}{216 \beta^2}
 \end{align*}
 Thus, \begin{align*}
 \frac{\partial \pi }{\partial \alpha} = \frac{\gamma q^{*2}(1-2\beta)}{72 \beta^2} \frac{\partial q^*}{\partial \alpha}
 \end{align*}
 Using the result of the Proposition \ref{thres} we finally obtain the result we were to prove.
\end{proof}
In contrast to the case with precise recommendation system (Proposition \ref{precise}), if the platform's precision is not very high (lower than 0.5), it may be optimal to charge maximal commission even in case when $\gamma > 1$. This happens since with very low precision the weight of quality for the recommendation becomes higher, as a result, sellers have incentives to reveal quality even with high commissions. In the similar fashion can be explained the fact that for low precision not very consumer-oriented recommender system (with $\gamma < 1$) prefers to set $\alpha = 0$: low precision adversely affects the consumer surplus, the situation deteriorates if in addition sellers decide not to disclose the information. Consequently, it is optimal for the platform to set $\alpha = 0$.
	\section{Discussion}
	\section{Extensions}
	\section{Conclusion}
	\newpage
	\bibliography{lib}{}
	%\bibliographystyle{unsrt}
	%\bibliographystyle{apalike}
	\bibliographystyle{abbrvnat}
\end{document}

>>>>

	One of the first papers which highlights the issue from the point of view of Economics is work by \cite{avery1999market}. \cite{avery1999market} considered the influence of the recommender system on choice regarding the consumption of good for the sequentially arriving agents. They model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review. The main focus of the work is on finding a socially optimal mechanism, in particular, the authors propose a mechanism which includes monetary transfers to some agents to induce them to try the good.

	\cite{bergemann2006optimal} address the question of optimal pricing under the presence of recommender system. They consider the seller with recommender system who is competing for unit mass of heterogeneous buyers with the competitive fringe of sellers who do not deploy the recommender system. The seller with recommender system uses reviews from past customers to make recommendations to the newcomers. The authors consider several type of equilibria concluding that the presence of the recommender system creates an additional value by reducing uncertainty for agents about the quality of the good.

@@

	One of the first works that addresses the question of optimal design of the recommendation policy is the work by \cite{kremer2014}. The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff. One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher. The finite number of consumers arrive sequentially and each chooses between either of two alternatives. After the choice is made the consumer leaves forever. The platform accumulates reviews and commits to the recommendation policy it uses to make recommendations to the newcomers. \cite{kremer2014}, firstly, neatly in simple terms demonstrate why the full revelation policy may be suboptimal\footnote{Maybe the appearance of this paper and similar to that in 2014 was the reason why \textit{Amazon} decided to switch their recommendation policy from the fully revealing one.}. Secondly, the optimal recommender policy is derived which has a threshold nature. That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold. Otherwise, the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end.

@@

	Another paper, which is closely related to the one mentioned previously is the paper by \cite{li2018recommender} (actually there are two quite different versions of this paper). In the first one, considering the same set-up as \cite{chen2016advertising}, the authors do not allow sellers to advertise themselves, assuming the shares of informed and uninformed customer to be exogenous. The authors demonstrate that the presence of the recommender system on one hand increase the share of informed customers and hence increases demand on goods, but on the other hand it intensifies competition by making a demand more elastic. \cite{li2018recommender} refer to these two effects as \textit{demand effect} and \textit{substitution effect} respectively. In the similar way as in \cite{chen2016advertising}, the authors demonstrate that the total effect of recommender system on sellers becomes a result of interaction between these tow effects. Also, in this paper the authors make the first attempt to endogenise the recommender policy, showing that the recommender system may increase its profit by changing the recommendation strategy. In the second version of the paper the authors elaborate on this, formulating the recommender system's objective and finding an optimal recommender policy the recommender system should implement. In the two (in fact, three) papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.

@@

\begin{proposition}\label{notprecise}
	Given the commission rate $\alpha$ and precision $\beta$ the optimal recommender policy implies:
	$$\hat{x} = \frac{1}{2} - \frac{\gamma - \alpha}{\gamma \beta} \frac{p_A-p_B}{2} + \frac{\tilde{q}_A - \tilde{q}_B}{2 \beta}$$
	i.e. those customers for whom signal regarding their location falls to the left of $\hat{x}$ are recommended the seller $A$, otherwise they are recommended the seller $B$.
	Equilibrium prices are
	$$p_i = \frac{\gamma}{\gamma - \alpha} \beta + \frac{\gamma(\tilde{q}_i - \tilde{q}_{-i})}{3(\gamma - \alpha)}$$
	where
	$$\tilde{q}_i = \begin{cases}
	q_i, \text{ if firm }i\text{ discloses }\\
	\frac{q^*}{2}, \text{otherwise}
	\end{cases}$$
	and \begin{align}\label{thres_b}
	q^* = -\frac{6 \beta - 1}{3} + \frac{1}{3 \gamma} \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}
	\end{align}
\end{proposition}
\begin{proof}
	Using \eqref{signal} and repeating the proof of the Proposition \ref{with} one can obtain the expression for $\hat{x}$ and equilibrium prices. $q^*$ is found in the same way as in the Proposition \ref{precise}.
\end{proof}
Note that for $\beta = 1$ i.e. for the highest precision of the recommender system, we get the exact result of the Proposition \ref{precise}. Several observations are worth noting from the Proposition above. First of all, low precision makes the impact of the recommender system less prominent. We have a multiplier $\frac{\gamma - \alpha}{2 \gamma \beta}$ before the term $(p_A - p_B)$ in the threshold which determines the demand for sellers. The lower is $\beta$ the stronger becomes incentives for the sellers to undercut the competitor. This can be explained in the following way: now there is a chance that the platform mistakes regarding the location of the customer, in this case, the platform makes a recommendation assuming that it is an average customer (location $\frac{1}{2}$). The latter increases the share of customers in the middle whom the sellers have the highest incentives to compete for. As a result, other things equal, equilibrium prices are lower, the less precise is the recommender system implemented by the platform.

@@

\begin{proposition}\label{profit_beta}
	\leavevmode 
	\begin{itemize}
		\item If $\beta \le 0.5$ and  $\gamma > 1$ then $\frac{\partial \pi }{\partial \alpha} > 0$ and the optimal commission is $\alpha = \bar{\alpha}$
		\item  If $\beta \le 0.5$ and $\gamma < 1$ then $\frac{\partial \pi }{\partial \alpha} < 0$ and the optimal commission is $\alpha = 0$	
		\item If $\beta > 0.5$ and  $\gamma > 1$ then $\frac{\partial \pi }{\partial \alpha} < 0$ and the optimal commission is $\alpha = 0$
		\item If $\beta > 0.5$ and $\gamma < 1$ then $\frac{\partial \pi }{\partial \alpha} > 0$ and the optimal commission is $\alpha = \bar{\alpha}$
	\end{itemize}
\end{proposition}
\begin{proof}
 First of all, we need to find the platform's expected profit. Given prices $p_A, p_B$ and qualities $q_A, q_B$, total benefit of the platform is equal to:
 \begin{align}\label{profit}
 \alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \biggl(\int_{0}^{\hat{x}} \biggl[ (\beta +(1-\beta)\hat{x})(V + q_A - p_A - x) + \nonumber \\+ (1-\beta)(1-\hat{x})(V + q_B - p_B - (1-x))\biggr]dx + \int_{\hat{x}}^1 \biggl[(1-\beta) \hat{x}(V+q_A-p_A-x) + \nonumber\\
 +(\beta + (1-\beta)(1-\hat{x}))(V + q_B - p_B -(1-x)) \biggr] dx\biggr) = \\
 = 	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(V - \frac{1}{2}+(1-\hat{x})(q_B - p_B) + \hat{x}(q_A-p_A) + \beta \hat{x} - \beta \hat{x}^2  \right)\nonumber
 \end{align}
 Where $\hat{x}$ is defined according to Proposition \ref{notprecise}. For those customers located to the left of $\hat{x}$ the probability of being recommended a seller $A$ is $\beta + (1-\beta) \hat{x}$. Since with probability $\beta$ the platform receives the true location of customer and with probability $(1-\beta) \hat{x}$ the platform receives a signal from the interval $[0, \hat{x}]$. In the following way, one can obtain probabilities of being recommended seller $B$ for those customers located to the left of $\hat{x}$ as well as probabilities of being recommended sellers $A$ and $B$ for the customers located to the right of $\hat{x}$. Finally, we obtain an expression in \eqref{profit}. Then, in the same way as in the Proposition \ref{opt_alpha} one can obtain the expression for the profit:
 \begin{align}\label{profit_b}
 \pi = \frac{\gamma(54 \beta^2(1+4V) - (1-q^{*3})(1-2\beta) - 216 \beta^3)}{216 \beta^2}
 \end{align}
 Thus, \begin{align*}
 \frac{\partial \pi }{\partial \alpha} = \frac{\gamma q^{*2}(1-2\beta)}{72 \beta^2} \frac{\partial q^*}{\partial \alpha}
 \end{align*}
 Using the result of the Proposition \ref{thres} we finally obtain the result we were to prove.
\end{proof}
In contrast to the case with precise recommendation system (Proposition \ref{precise}), if the platform's precision is not very high (lower than 0.5), it may be optimal to charge maximal commission even in case when $\gamma > 1$. This happens since with very low precision the weight of quality for the recommendation becomes higher, as a result, sellers have incentives to reveal quality even with high commissions. In the similar fashion can be explained the fact that for low precision not very consumer-oriented recommender system (with $\gamma < 1$) prefers to set $\alpha = 0$: low precision adversely affects the consumer surplus, the situation deteriorates if in addition sellers decide not to disclose the information. Consequently, it is optimal for the platform to set $\alpha = 0$.

In the exposition above I treated the precision of the recommender system $\beta$ as given, since I think it may be a reasonable assumption for the platform in the short-run. In real life, this precision is usually determined by the amount of data the platform collected about customers as well as the complexity of algorithm the platform employs to process and use this data. And once the amount of data and the algorithm are chosen and  fixed, the precision cannot be improved in the short-run and, hence, may be supposed to be fixed. At the same time, the platform may invest resources into increasing the precision in the long-run. And the question that logically arises is: is it also beneficial for the platform to increase the precision of the recommender system? Interestingly, the model I consider suggests that it might not always to be the case.
\begin{proposition}
	Given $\alpha$, the perfect precision $\beta = 1$ is always worse for the platform than some intermediate $\beta \in (0, 1)$
\end{proposition}
\begin{proof}
	Plugging \eqref{thres_b} into \eqref{profit_b} one can obtain
	\begin{align*}
	\pi = \frac{\gamma}{216 \beta ^2}  \biggl\{-2 \beta  \left(\frac{\left(\sqrt{\gamma  \left(\frac{216 \beta  c (\gamma-\alpha)}{1-\alpha}+(1-6
				\beta )^2 \gamma \right)}-6 \beta  \gamma +\gamma \right)^3}{27 \gamma ^3}-1\right)+\\
			+\frac{\left(\sqrt{\gamma 
				\left(\frac{216 \beta  c (\gamma - \alpha)}{1-\alpha}+(1-6 \beta )^2 \gamma \right)}-6 \beta  \gamma +\gamma
			\right)^3}{27 \gamma ^3}-216 \beta ^3+54 \beta ^2 (4 V+1)-1\biggr\}
	\end{align*}
	Then
	\begin{align*}
	\pi\bigg|_{\beta = 1} = \frac{\gamma }{216} \biggl\{\frac{\left(\sqrt{\gamma  \left(\frac{216 c (\gamma-\alpha)}{1-\alpha}+25 \gamma \right)}-5
		\gamma \right)^3}{27 \gamma ^3}-2 \left(\frac{\left(\sqrt{\gamma  \left(\frac{216 c (\gamma-\alpha)}{1-\alpha}+25
			\gamma \right)}-5 \gamma \right)^3}{27 \gamma ^3}-1\right)+\\
		+54 (4 V+1)-217\biggr\}
	\end{align*}
	At the same time, 
	\begin{align*}
	\pi \bigg|_{\beta = \frac{1}{2}} = \gamma  \left(V-\frac{1}{4}\right)
	\end{align*}
	Consider the difference $\pi\bigg|_{\beta = 1} - \pi\bigg|_{\beta = \frac{1}{2}}$:
	\begin{align*}
	\pi\bigg|_{\beta = 1} - \pi\bigg|_{\beta = \frac{1}{2}} = -\frac{\gamma}{216} \biggl\{ 107+ \frac{\left(\sqrt{\gamma  \left(\frac{216 c (\gamma-\alpha)}{1-\alpha}+25 \gamma \right)}-5
		\gamma \right)^3}{27 \gamma ^3} \biggr\} < 0
	\end{align*}
	Thus, setting $\beta = \frac{1}{2}$ does the platform strictly better-off comparing to the case $\beta = 1$.
\end{proof}
	\section{Discussion}
	\section{Extensions}
	\section{Conclusion}
	\newpage
	\bibliography{lib}{}
	%\bibliographystyle{unsrt}
	%\bibliographystyle{apalike}
	\bibliographystyle{abbrvnat}
\end{document}
